Supporting Information : Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes

### 2018-11-19 Ben_Manuscripts/structure_paper/response_to_reviewers.tex

    We have modified the paragraph of our conclusion which describes future
    work (page~\pageref{M-addition:forcefield-structure}) in order to acknowledge
    the possibility that a more tuned force field could influence our results:

@@

    Old text: ``Future work, based on what has been learned in this study, may
    help further improve the structural agreement between experiment and
    simulation"

    New text: ``Future work, based on what has been learned in this study,
    may help further improve the structural agreement between experiment and
    simulation and test the sensitivity of our current conclusions to a more
    accurate structure."

@@

    \item \begin{quote}
    \textit{The MD-Structure-Factor repository is not available at the provided
	    github link, nor does it appear to be online anywhere. So, the correctness of
	    that code cannot be checked.}
    \end{quote}

    Author reply: We thank the reviewer for pointing this out and apologize for
    our mistake. The referenced GitHub repository was set to private due to
    miscommunication with our collaborators but it has now been made public. Please
    feel free to read through the code. We have performed checks to ensure that the
    code works properly, especially when simulating monoclinic unit cells. For
    example, we generated test trajectories containing thousands of frames made up
    of configurations of randomly placed point scatterers in a monoclinic unit cell.
    We expect that the simulated diffraction pattern will have a gaussian shape.
    This exact result is shown in the Figure~\ref{S-fig:xrd_noise} of the SI.
    With improper treatment of the monoclinic geometry, the gaussian shape is
    skewed into an elliptical shape.

@@

    \item \begin{quote}
    \textit{One might want to check that code based on Figure 7f, where
	    presumably spherical scatterers generate reflections with anisotropy (diamonds
	    instead of squares or circles). This kind of an artefact is plausible if there
	    are errors handling the non-cubic box, or with scattering angles that aren't
	    parallel to a box axis. llcsim/analysis/correlation.py looks like it has the
	    machinery for doing the scattering analysis.}
    \end{quote}

    Author reply: We thank the reviewer for their astute observation and
    the resulting improvements to our manuscript. We generated Figure 7f by
    calculating the structure factor of a 2-dimensional array of hexagonally-packed
    scatterers. It is meant to be a simplified representation of the scattering
    exhibited by hexagonally packed alkane tails so that we can qualitatively
    understand where we would expect R-spots to appear if tails indeed pack in that
    way. Since it is possible to generate a rectangular unit cell which contains
    hexagonally packed point scatterers, there are no artifacts due to improper
    treatment of periodicity.  Although we believe our calculations are correct, we
    originally used only eight total scatterers in the 2D unit cell and then
    plotted the structure factor with a contour plot
    (\texttt{matplotlib.pyplot.contourf}) which does not interpolate the data. We
    re-created Figure 7f by simulating the structure factor of 100 hexagonally
    packed scatterers, in order to get higher fourier space resolution, and then
    plotting the result with \texttt{matplotlib.pyplot.imshow} and Gaussian
    interpolation. We have added the code used to generate this figure to the
    \texttt{llcsim} repository which can be accessed at
    \texttt{llcsim/analysis/hexagonal\_packing.py}.
    Additionally, we have added an entry to Table~\ref{S-table:python_scripts} in 
    order to make it clear where the code can be found.

@@

    \item \begin{quote}	
    \textit{Regarding the ion conductivities, how do the numbers presented here
	    compare against ion conductivities predicted with GAFF and other polarizable
	    and nonpolarizable force fields? If GAFF is systematically too high in other
	    cases, it could help explain the order-of-magnitude discrepancy observed here.}
    \end{quote}

	Author reply: We thank the reviewer for bringing up a good point. There are few studies
	with experimental comparisons of sodium ion conduction. Conduction of pottasium in 
	ion channels is quite common. The ionic conductivity is directly related to the 
	diffusivity of the ion of interest, so if the dynamics of sodium are overpredicted
	in our model, this could give rise to the discrepancy between experiment and simulation.
	The diffusivity of sodium ions has been overpredicted with multiple models, including AMBER
	where the diffusivity is about 2x too high. We have modified the text on page
	\pageref{M-modification:conductivity} to include additional reasoning for the 
	discrepancy between the simulated and experimental ionic conductivity.

	Old text: ``The calculated value of ionic conductivity is 5 times higher than experiment 
    likely because we simulated infinitely long, aligned pores. The ionic conductivity 
    measurement to which we are comparing was done with an \SI{80}{\micro\metre}-thick film, 
    nearly 10,000 times thicker than our simulated system. The thick film is likely 
    imperfectly aligned and has defects leading to non-contiguous pores. It has been shown
    that there is a large dependence of ionic conductivity on the alignment of the pores.
    The ionic conductivity of an isotropically aligned film is ca. 85 times lower than that
    of the nearly aligned film to which we are comparing.$^{16}$ We 
    hypothesize that a thin, perfectly aligned film would have a value of ionic conductivity
    in closer agreement with our model."

    Revised text: ``The calculated value of ionic conductivity is 5 times higher than
    experiment likely because we simulated infinitely long, aligned pores {\color{red} and
    because our model over-predicts the diffusivity of sodium}. The ionic conductivity 
    measurement to which we are comparing was done with an \SI{80}{\micro\metre}-thick film, 
    nearly 10,000 times thicker than our simulated system. The thick film is likely 
    imperfectly aligned and has defects leading to non-contiguous pores. It has been shown
    that there is a large dependence of ionic conductivity on the alignment of the pores.
    The ionic conductivity of an isotropically aligned film is ca. 85 times lower than that
    of the nearly aligned film to which we are comparing$^{16}$. 
    {\color{red} Additionally, sodium ions parameterized with AMBER parameters exhibit bulk
    water diffusion coefficients that are two times greater than experiment.$^{56}$ We cannot 
    definitively say how this ratio changes in the confined LLC pore environment, but
    it is likely that sodium diffuses faster in our system which contributes to a higher 
    ionic conductivity than experiment.}"

	% Amber parameters used for sodium
	% most studies with conductivity look at potassium
	% look at diffusivities

@@

	\item \begin{quote}
    \textit{If the authors either (a) strengthen the cases for their proposed structural
    insights being  highly likely (not just plausible), or (b) make a stronger case for
    why these particular insights have important ramifications for our understanding of 
    LLC's (their significance), then it would be easier to publish this immediately. This
    reviewer's opinion is that this work would find a broader interested audience with a more
    concise focus on these physical insights, which could be accomplished by moving more
    content (e.g., dihedral correlations) to the SI.}
    \end{quote}

    Author reply: We feel that the additions and modifications to the sections cited above 
    have made the intent of our work clearer and helped communicate the significance of 
    our learned insights for future work on this topic. We have clarified the purpose of work
    and what we expect to learn from it. We have commented on the significance of knowing the
    number of columns-per-pore. We have emphasized the observation that the radial distribution
    of monomer components is relatively independent of the system's initial configuration
    which is important for future high-throughput work. We have clarified and strengthened our
    reasoning for the existence of the R-double feature.

    We sympathize with the reviewer's desire for a more concise manuscript but feel that most 
    of the information presented is necessary in order to come to a full understanding of the 
    work we've performed (an opinion shared by Reviewer 2). As a compromise, we've made efforts 
    to shorten the manuscript by moving figures and discussions which draw the reader's attention
    from the important physical insights to the SI. Excluding additional citations (new citations
    take up about a page in the new draft), we have reduced the total length of the manuscript 
    by 3 pages (including any additions documented above). The following changes made this possible:

    \begin{itemize}

    \item We moved the plots of the 1D Correlation functions, $g(z)$, used to calculate 
    correlation lengths from Section~\ref{M-section:rpi} to the SI, now Figure~\ref{S-fig:correlation}.
    The most important information, the correlation lengths, is already presented in 
    Table~\ref{M-table:correlation_length}.

    \item We moved the discussion of the influence of system size on $g(z)$ 
    from Section~\ref{M-section:rpi} to the Supporting Information, 
    Section~\ref{S-section:correlation_functions}. We supplemented that discussion in
    the main text with the following text:

    ``We show that the system size in the $z$-direction does not significantly 
    alter $g(z)$ in Section~\ref{S-section:correlation_functions} of the SI."

    \item We removed the discussion of a system that we annealed from 500 K to 300 K
    from Section~\ref{M-section:rpi} because it is described again (with better context)
    in Section~\ref{M-section:slow_dynamics}.

	\item We moved the discussion of a correlation function, $g(z)$, that averages
	all slices of the 3D correlation function and the associated figure from 
	Section~\ref{M-section:rpi} to Section~\ref{S-section:correlation_functions}.

    \item We've supplemented the discussion of the influence of increased noise in the 
    $z$-direction and in the $xy$ plane on the shape and intensity of R-$\pi$ with the
    following paragraph:

    ``We also used our simplified systems in order to study the influence of increased
    disorder in each dimension. Increasing $z$-directional disorder reduces the
    intensity of R-$\pi$ without changing the shape of its $q_r$ cross-section.
    Increasing disorder on the $xy$ plane, somewhat counterintuitively, reduces
    the FWHM of the $q_r$ cross-section of R-$\pi$ with an insignificant effect
    on its intensity. We explain these points in more depth in 
    Section~\ref{S-section:simplified_sf}."

    The full discussion, with figures, has been moved from Section~\ref{M-section:rpi}
    to Section~\ref{S-section:simplified_sf}. The subsection of Section~\ref{M-section:rpi}
    starting on page~\pageref{M-section:rpi_xsection} of the main text now has a primary
    focus on the relationship between R-$\pi$ and inter-column correlation.

    \item We moved the plots of dihedral autocorrelation functions to the SI, 
    Figure~\ref{S-fig:dihedral}.

    \item We moved the figure associated with physical property changes after cross-linking
    to the SI, Figure~\ref{S-fig:xlink}, since its main findings are summarized sufficiently
    in the text.

    \end{itemize}

@@

    Author reply: We thank the reviewer for pointing out these minor errors in our main text. These issues have been corrected in
    the revised main text file. % need to add page numbers once finished

    \item \begin{quote} \textit{(3) Page 54: mention examples of force fields that explicitly include pi-pi interactions
    } \end{quote}

    Author reply: We have adjusted our language in order to specifically mention and cite polarizable force fields as follows:

    ``Polarizable force fields such as AMOEBA which explicitly including $\pi-\pi$ interactions
    between aromatic functional groups may be able to draw stacked monomers closer together. 
    $^{57-59}$"

    We have added citations with relevant force field examples, including AMOEBA:

    \begin{itemize}

    \item Shi, Y.; Xia, Z.; Zhang, J.; Best, R.; Wu, C.; Ponder, J. W.; Ren, P. Polarizable Atomic
    Multipole-Based AMOEBA Force Field for Proteins. J. Chem. Theory Comput. 2013,
    9, 4046–4063.

    \item Kaminski, G. A.; Stern, H. A.; Berne, B. J.; Friesner, R. A. Development of an Accu-
    rate and Robust Polarizable Molecular Mechanics Force Field from ab Initio Quantum
    Chemistry. J. Phys. Chem. A 2004, 108, 621–627.

    \item Lopes, P. E. M.; Lamoureux, G.; Roux, B.; MacKerell, A. D. Polarizable Empirical
    Force Field for Aromatic Compounds Based on the Classical Drude Oscillator. J. Phys.
    Chem. B 2007, 111, 2873–2885.

    \end{itemize}
